%
% Conclusion
%
\section{Conclusion}
\label{sec:symdynconclusion}
In this article we generalize symbolic dynamics for chaotic maps and define a new object called
stochastic symbol sequence. For a set of probability distributions in the reduced space
$\bar{\Omega}$, they can be evolved easily by a shift operator, and the total variation distance to
the invariant distribution can be bounded by the distance between two binomial distributions. We
show that for the distribution $\omega^0$ with stochastic symbol sequence $\delta_i=1/2+\epsilon
r^i$, $r \to 1$, its $|\omega^k-\bar{\omega}|_{TV}$ has normal shape. Thus we can generate
a sequence of $\omega^0_n$ to produce a sequence of cutoff trajectories that have the same
limiting behavior as the ones found in the random walk on an $n$-dimensional hypercube problem. This
result can be generalized to other cutoffs with normal shape in finite Markov Chains.

The above result not only demonstrates the close relationship between cutoff phenomenon studies in finite Markov Chain and chaotic maps, but also points out the source of cutoffs: almost independence of the underlying small processes. This can be very useful for the study of large systems. So far there are not many good tools to understand and analyze very large systems. However, if we can identify each of the small processes in the large system and justify their almost independence, cutoffs can be proved. 

The relation between chaos and cutoff phenomenon has also been found in numerical studies and
experiments. In \cite{numcutoff}, the author provides numerical evidence to show that the
variance of Standard Map presents a cutoff when the diffusion goes to zero. In \cite{topopt} a
numerical simulation of a chaotic mixing channel also shows cutoffs in mixing trajectories. More
numerical evidence can be found in \cite{Thiffeault2003-13, Tsang2005}, though they do not call
the phenomenon cutoff. The above studies focus on the convergence trajectories of the variance of a
scalar function advected by 2-D chaotic maps with small diffusion---a different content from the study here. However, just as it is believed that cutoff phenomenon is widespread in many
finite Markov Chains, we believe cutoffs are intrinsic for chaotic maps, not only for the
distribution inside or close to $\bar{\Omega}$.
